#440 Reproducibility and Statistics Assignment Example

<!-- This is an Rmd comment. -->
<!--
You can make comments several lines long by inserting
the text in between the the start and the end of the 
comment.

Comments are where I'll tell you what to do. You should
use the space directly below the comments to do what
the comment asks. You should also describe what you're
doing to make it obvious to the reader so they can
easily follow your logic in a fully reproducible way.
For any statistical tests you run, you should explain
the results in plain English and present the statistics
as you would in an APA style manuscript. For any symbols
you don't know how to use/can't use, feel free to use the name
instead (e.g. alpha)
--> 

<!-- Load the politics.csv data file. -->
First we need to load the data.

```{r}
pols<-read.csv("politics.csv")
```

<!-- 
Check the structure.
subject, party, testtime, minwage, and sex need to be factors.
optimismscore and income should be integer and numeric vectors.
Fix any variables that aren't right.
-->
Let's take a look at the data.
```{r}
str(pols)
```

Subject should be a factor variable. Additionally, I prefer
it when "pre" precedes "post". So I'm going to refactor those
variables.
```{r}
pols$subject<-factor(pols$subject)
pols$testtime<-factor(pols$testtime, levels=c("pre", "post"))
```

<!-- Find the minimum, mean, and median incomes. -->
I'd like to find the minimum, mean, and median incomes. I can do
all this and more in one fell swoop.

```{r}
summary(pols$income)
```

<!-- Create a histogram of incomes.
Feel free to use the default ugly settings since
we didn't cover how to do this in ggplot2. If
you're a real go-getter, see if you can figure
out how to make a prettier histogram with better
labels, etc. using ggplot2. -->
Let's see how incomes are distributed. I noticed that there
each income is represented twice. So to avoid over counting
incomes, I'll only use the data from the posttests to create
a histogram.

```{r}
hist(pols$income[pols$testtime=="post"])
```

<!-- Run a test to determine whether political affiliation
is independent of one's support to increase the minimum wage before
they watch the videos.
Present the data supporting your analyses in graphical or
tabular form.
-->

Let's see if political affiliation and minimum wage support
are independent before the participants watched *America's Funniest
Home Videos*. To do this I need to calculate the frequency
that individuals appear in different political groups. Then I'll
use a Chi-Square test of independence to test the hypothesis that
affiliation and support are independent. Again, I don't want
to count people twice, so I'll focus on the posttest data.

```{r}
tab<-table(pols$party[pols$testtime=="pre"],pols$minwage[pols$testtime=="pre"])
tab
```

These are the frequencies individuals fell into the different
conditions.

```{r}
chisq.test(pols$party[pols$testtime=="pre"],pols$minwage[pols$testtime=="pre"])
```

Political affiliation is not independent of support for minimum
wage increases, Chi-Square(2) = 6.4, p = .04.

<!-- Run a test to determine whether or not males and females
have different optimism scores before they watch the videos.
Create a graph to display your data that includes the relevant
means and standard errors.
Note we didn't create this graph in the instructions. We're doing
this instead of the first bar graph in the instructions.
-->
To see whether or not males and females have different optimism
scores before watching *AFHV* we can run an independent t-test.

```{r}
t.test(pols$optimism[pols$sex=="male" & pols$testtime=="pre"],pols$optimism[pols$sex=="female" & pols$testtime=="pre"])
```

Males and females do not differ in optimism before watching *AFHV*,
t(62.9) = -.37, p = .71.

To create the figure, I need to include the dplyr library. Then
I'll use it to summarize the politics data set and find the means
and standard errors for the different conditions. Then I'll need
the ggplot2 library to create a nice plot that's easy to work with.
I'll do the same thing for any other graphs I create unless I
specify otherwise.

```{r}
library("dplyr")
temp<-pols[pols$testtime=="pre",]%>%group_by(sex)%>%summarize(means=mean(optimismscore),
              sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("ggplot2")
f<-ggplot(temp,aes(x=factor(sex),y=means))+
    geom_bar(stat="identity", color="black",fill=c("deeppink","deepskyblue2"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems), width=.1)
f
```

That's a bare minimum graph that might be ok for others, but not for me.

```{r}
f<-f+ggtitle("Video Optimism Scores by Sex")+
    labs(x="Sex", y="Pretest Optimism Scores\n(higher=more optimistic")+
    scale_x_discrete(breaks=c("female","male"),labels=c("Female","Male"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(), axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())
f
```

That's much better.

<!-- See if male and females with different political affiliations
have different incomes. Create a graph to display your data that
includes the relevant means and standard errors.-->
Let's see if party affiliation and sex predict incomes independently
or in an interaction using a 2-way between-subjects ANOVA.

```{r}
 summary(aov(income~party*sex,data=pols[pols$testtime=="pre",]))
```

Income is related to party affiliation, F(2, 60) = 4.54, p = .015, but not
to sex, F(1, 60) = 3.58, p = .064, or an interaction between affiliation
and sex, F(2, 60) = .07, p = .93.

To create the plot and manually adjust the fill, I need to use
hexadecimal color codes. To convert the named colors I like
to hexadecimal, I need to use the gplots library.

```{r}
temp<-pols[pols$testtime=="pre",]%>%group_by(party,sex)%>%
    summarize(means=mean(income),sems=sd(income)/sqrt(length(income)))
library("gplots")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, fill=sex))+
    geom_bar(stat="identity",position=position_dodge())+
    scale_fill_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    theme(legend.key=element_rect(color="black"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2,position=position_dodge(.9))+
    ggtitle("Incomes by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Income\n(thousands of dollars)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

<!-- Create a line graph for the same data. With
error bars of course!-->

Although, we shouldn't make a line graph for these data, my
moronic instructor thinks I need the practice. Oh, shoot. I
hope he doesn't dock my grade for that. Whatevs. lol! :)

<!-- Don't use slang, 1377 speak, emoji, or abbreviations
that are so popular in texting. Lulz ;) is not ok for this
assignment.-->

```{r}
f<-ggplot(temp, aes(x=party, y=means, group=sex, color=sex))+
    geom_line(size=1)+
    geom_point(size=2)+
    scale_color_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
    ggtitle("Incomes by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Income\n(thousands of dollars)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

<!-- Run a test to see if optimism scores pre- and post
watching videos depends on party affiliation-->

To see if optimism scores before and after watching *AFHV*
vary depending on party affiliation, I'm going to run a
2-way mixed ANOVA. I'm using a mixed ANOVA because the same
people indicated their optimism before and after the videos
(i.e. a within-subjects factor) and because different
people are affiliated with different political parties (i.e.
a between-subjects factor).

```{r}
summary(aov(optimismscore~testtime*party+Error(subject/testtime),data=pols))
```

Although optimism scores were related to political affiliation, F(2, 63)
= 32.56, p < .001, and affected by watching videos, F(1, 63) = 42.53, p
< .001, there was no interaction between taking the test before or after
watching videos and their party affiliation, F(2, 63) = 1.48, p = .24.

<!-- Run a Multiple Regression Analysis to see
whether you can predict someones pretest optimism
scores from their income and sex. Create a figure
that shows the distributions of pretest optimism,
income and sex and create lines that best fit the
data for both sexes regardless of whether the lines
are warranted based on the statistical analyses-->

I'm going to perform Multiple Regression
to predict pretest optimism scores from income and
sex.

```{r}
summary(lm(optimismscore~income+sex,data=pols[pols$testtime=="pre",]))
```
The linear model predicting pretest optimism scores from income
and sex doesn't explain any variance, R^2 = .014, F(2, 63) = .44,
p = .65. Despite this lack of significance, I'm creating a
scatter plot because my wonderful instructor thinks that practice
makes perfect.

```{r}
f<-ggplot(pols[pols$testtime=="pre",],aes(x=income,y=optimismscore,color=sex))+
    geom_point(size=2)+
    geom_abline(intercept=60.2+2.51/2, slope=-.092,color=col1)+
    geom_abline(intercept=60.2-2.51/2, slope=-.092,color=col2)+
    scale_color_manual(values=c(col1,col2),breaks=c("female","male"),labels=c("Female","Male"))+
    ggtitle("Optimism Predicted by Sex and Income")+
    labs(x="Income (Thousands of Dollars)",y="Optimism Score\n(Higher=More)")+
    theme(plot.title=element_text(size=15,face="bold", vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold", vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold", vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    theme(panel.border=element_blank(), axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.minor.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.83,.86))+
    theme(legend.background=element_blank())+
    theme(legend.title=element_blank())+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

***Fin***