#How the data were generated

The first thing I did was to set a random seed to ensure that the data
are fully reproducible.

```{r}
set.seed(2073) #I could have used any number. 2073 is just the first I though of.
```

Next, I determined how many participants (N) I would have for Republicans (rN),
Independents (iN), and Democrats (dN). Imagine participants had to choose which
party they identified with the most. I made sure I had an even number so
I could have equal numbers of males and females. I also determined the total
number of participants so I could generate the rest of the data more easily.

```{r}
rN<-23
iN<-17
dN<-26
#Make sure the total is even
totN<-rN+iN+dN
```
I created party affiliation by repeating each type of participant and combining them
together. I then doubled these affiliations so I could have two observations of each
participant.

```{r}
party<-rep(c(rep("republican",rN),rep("independent",iN),rep("democrat",dN)),times=2)
```
Then I created a variable to indicate whether the observation came before or after
my crucial imaginary manipulation (remember, these are completely fabricated data).
```{r}
testtime<-rep(c("pre","post"),each=totN)
```
Next, I created a variable for optimism scores. I imagined I had some survey with 10
10-point Likert scale questions. The scores here, represent the total score that could
range from 0 (not at all optimistic) to 100 (completely optimistic). I created the
pretest scores by sampling from a random normal distribution in a way that should
correlate with affiliation. To create the posttest scores, I added values from
another random normal distribution to the original scores so the scores should
correlate.
```{r}
optimismscore<-round(rnorm(totN,30+50*(1:totN)/totN, 12))
optimismscore<-c(optimismscore, optimismscore+round(rnorm(totN,5,7)))
```
Next, I created a binary variable to indicate agreement with the imaginary
statement "The minimum wage should be increased" before and after the manipulation.
I created this variable so it should correlate with political affiliation and so
the pre and post scores would correlate with one another. Then I changed the
variable to a factor with levels no and yes.
```{r}
minwage<-c(rbinom(totN,1,.2+.6*(1:totN)/totN),rbinom(totN,1,.3+.55*(1:totN)/totN))
minwage<-as.factor(minwage)
levels(minwage)=c("no", "yes")
```
Next, I created another binary variable to indicate the sex of the participant.
```{r}
sex<-rep(sample(rep(c("male","female"),each=totN/2)),times=2)
```
Then, I generated incomes for each participant based on their political affiliation.
```{r}
income=rep(c(rweibull(rN,1.8,54),rweibull(iN,3,56),rweibull(dN,2.2,32)),times=2)
```
I didn't want the data to be grouped by party or be roughly in order of optimism
scores, as that would almost never happen in the real world when random sampling.
To mix things up, I randomly sampled the total number of participants to create
a mixed ordering that would keep data for each participant together.
```{r}
order<-sample(1:totN)
order<-c(order,order+totN)
```
Then I created a variable for each subject that indicates whether they were first,
last, or somewhere in the middle.
```{r}
subject<-as.factor(rep(1:(totN),times=2))
```
With all the data I need to run the analyses my students are likely to encounter,
I stitched the variables together as a data frame before saving the data frame
as politics.csv.
```{r}
politics<-data.frame(subject=subject,
                    party=party[order],
                    testtime=testtime[order],
                    optimismscore=optimismscore[order],
                    minwage=minwage[order],
                    sex=sex[order],
                    income=income[order])
write.csv(politics, "politics.csv", row.names=FALSE)
```