---
output: html_document
---
#440 Reproducibility and Statistics Assignment Homework

<!-- Veronica Conaway 
Reproducibility Homework. -->
<!--
-->I must set the working directory in order to acess the information.
```{r}
setwd("~/Research/reproduce")
```

<-Here I am going to load the data file that I will be using.->
```{r}
pols<-read.csv("politics.csv")
```

<- I will now check the structure of the data which will give me the subject, party, testtime, minwage, and sex.->
```{r}  
str(pols)
```
<-Here I need to assign the factor variable and also make sure that the variable testtime is defined from pre and post test.->
```{r}
pols$subject<-factor(pols$subject)
pols$testtime<-factor(pols$testtime,levels=c("pre","post"))

```
<- I am running a code to find the minimum, mean, and variance for posttest optimism scores.->
```{r}
summary(pols$optimismscore[pols$testtime=="post"])
var(pols$optimismscore[pols$testtime=="post"])
```
The min is 18, the mean 59.82, and the variance for the posttest optimism score is 336.27

<-The histogram that I have made shows optimism scores.->
```{r}
hist(pols$optimismscore[pols$testtime=="post"])
```

<!-- In order to determine if political affiliation
is independent of one's gender I have created a table with from the data and then put into the Pearson's Chi-Squared test.-->
```{r}
tab<-table(pols$party[pols$testtime=="pre"],pols$sex[pols$testtime=="pre"])
chisq.test(pols$party[pols$testtime=="pre"],pols$sex[pols$testtime=="pre"])
```
  The Chi-Square(2)=0.73, p = 0.70 this shows that political affiliation is not independent to gender.  
  
<!-- A Welch two sample t.test is run to determine whether or not males and females have different incomes.
-->
```{r}
t.test(pols$income[pols$sex=="male" & pols$testtime=="pre"],pols$income[pols$sex=="female" & pols$testtime=="pre"])                
```
Income does not differ based on sex t(61.62)=-1.57,p=0.12


I did not have these on my Rstudio so in order to move on I had to install the items to continue. 
```{r}

install.packages("dplyr")
install.packages("ggplot2")
install.packages("gplots")
library("dplyr")
library("gplots")
library("ggplot2")
```
<---Here is the graph that shows the diffrence in income based on sex.--->  
```{r}  
temp<-pols[pols$testtime=="pre",]%>%group_by(sex)%>%summarize(means=mean(optimismscore),sems=sd(income)/sqrt(length(income)))
f<-ggplot(temp,aes(x=factor(sex),y=means))+geom_bar(stat="identity",color="black",fill=c("deeppink","deepskyblue2"))+geom_errorbar(aes(ymax=means+sems, ymin=means-sems), width=.1)
f+ggtitle("Incomes by Sex")+labs(x="Sex", y="Incomes\n(higher=moreincome")+scale_x_discrete(breaks=c("female","male"),labels=c("Female","Male"))+theme(plot.title=element_text(size=15,face="bold",vjust=.5))+theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+theme(axis.text.x=element_text(size=10,face="bold",color="black"))+theme(axis.text.y=element_text(size=10,face="bold",color="black"))+coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+theme(panel.border=element_blank(),axis.line=element_line())+theme(panel.grid.major.x=element_blank())+theme(panel.grid.major.y=element_line(color="darkgrey"))+theme(panel.grid.minor.y=element_blank())
```
<!-- Anova is ran to see male and females with different political affiliations have different posttest optimism scores.
```{r}
summary(aov(optimismscore~party*sex,data=pols[pols$testtime=="post",]))
```
Based on the Anova post optimism scores are related to party, F(2,60)=27.06, p<0.001, but sex was not related to post optimism scores F(1,60)=0.04,p=0.84.  Also there was no relation between sex and party on post optimisim scores F(2,60)= 1.21,p=0.30.
```{r}
temp<-politics[politics$testtime=="post",]%>%group_by(party,sex)%>%
    summarize(means=mean(optimismscore),sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("gplots")
library("ggplot2")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, fill=sex))+
    geom_bar(stat="identity",position=position_dodge())+
    scale_fill_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    theme(legend.key=element_rect(color="black"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2,position=position_dodge(.9))+
    ggtitle("Post-Test Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Post-Test Optimism\n(Higher Value=Greater Optimism)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```
 
This shows that scores are related to party but not sex or a combanation of both.  

```{r}

temp<-pols[pols$testtime=="post",]%>%group_by(party,sex)%>%
    summarize(means=mean(optimismscore),sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("gplots")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, fill=sex))+
    geom_bar(stat="identity",position=position_dodge())+
    scale_fill_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    theme(legend.key=element_rect(color="black"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2,position=position_dodge(.9))+
    ggtitle("Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Optimism Scores)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.7,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```
<!-- Here is a line graph for the same data-->
```{r}
f<-ggplot(temp, aes(x=party, y=means, group=sex, color=sex))+
    geom_line(size=1)+
    geom_point(size=2)+
    scale_color_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
    ggtitle("Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Optimism Scores)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```
<!-- A Anova test was run to see if there where any relation of scores and sex and testtiime.  It is shown that there is no relation between sex F(1,64)=0.12,p=0.73, and scores but there is a relation in watching videos F(1,64)=41.30,p<0.001 but it did not matter when the video was watched F(1,64)=0.05,p=0.83. -->
```{r}
summary(lm(optimismscore~income+sex,data=pols[pols$testtime=="pre",]))
```

<!-- R2=0.90, F(3,62)=180.3,p<0.001 shows that pretest optimism scores predict posttest optimism scores with a slope of 0.94.     
```